Tulai

### 2023-09-25 paper/paper.qmd
changed calculation for 'total_lithics' and "core_count"

tl4 <- tl4 %>%
  mutate(
    `Core Typology` = str_to_lower(`Core Typology`),
    `Core Typology` = str_trim(`Core Typology`),
    # BM: what does this regex do in the next line?
    `Core Typology` = str_replace_all(`Core Typology`, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(`Core Typology` = case_when(
    `Core Typology` == "flatcore" ~ "flat-pressure",
             `Core Typology` == "rejuvention" ~ "rejuvenation piece-NA",
             `Core Typology` == "pressureprymidal core" ~ "pyramid-pressure",
             `Core Typology` == "pressureflat core" ~ "flat-pressure",
             `Core Typology` == "pressurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prussurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prismaticcore" ~ "prismatic-percussion",
             `Core Typology` == "pyramidalpressure core" ~ "pyramid-pressure",
             `Core Typology` == "flatcore with one debitage surface/pressure" ~ "flat-pressure",
             `Core Typology` == "corefragment" ~ "core fragment-NA",
             `Core Typology` == "cilandrical/pressure" ~ "cylinder-pressure",
             `Core Typology` == "bullet/pressure" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/pressure" ~ "pyramid-pressure",
             `Core Typology` == "percussion/pyramidal" ~ "pyramid-pressure",
             `Core Typology` == "flat/pressure" ~ "flat-pressure",
             `Core Typology` == "pyramidal/ percussion??" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ peressure" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ pressure" ~ "pyramid-pressure",
             `Core Typology` == "pressure" ~ "pyramid-pressure",
             `Core Typology` == "multidirectional/percussion" ~ "shapeless-percussion",
             `Core Typology` == "prismatic/ percussion" ~ "pyramid-pressure",
             `Core Typology` == "flatcore/ unidirectional/pressure" ~ "flat-pressure",
             `Core Typology` == "heavilyused" ~ "NA",
             `Core Typology` == "burnt" ~ "NA",
             `Core Typology` == "patinated" ~ "NA",
             `Core Typology` == "pyramidal(bullet)/pressure" ~ "bullet-pressure",
             `Core Typology` == "cylandrycal/ bidirectional pressure core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure bullet core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "bladelet" ~ "bullet-pressure",
             `Core Typology` == "lip" ~ "NA",
             `Core Typology` == "bladlet" ~ "bullet-pressure",
             `Core Typology` == "bladelet(30.56)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(28.55.6)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(16.28.1)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(40 5.7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(23.2 2.3)" ~ "bullet-pressure",
             `Core Typology` == "bladelet-flakeblade(41 16)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(3512.6/7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(27.307.5)" ~ "bullet-pressure",
    `Core Typology` == "pressure/pyramidal" ~ "pyramid-pressure",
    `Core Typology` == "bladeletcore" ~ "bullet-pressure",
    `Core Typology` == "pyramidal? / pressure" ~ "pyramid-pressure",
    `Core Typology` == "prismatic" ~ "pyramid-pressure",

    `Core Typology` == "bullet" ~ "bullet-pressure",

@@

@@

```{r}
tl_final <- tl_final %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`))  %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`)) %>%
  mutate(
    `core-fragment` = case_when(
      `core-fragment` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-fragment` == "rejuvention piece-side" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece-platform" ~ "core tablet",
      `core-fragment` == "core preparation" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece" ~ "rejuvenation piece",
      `core-fragment` == "core preperation" ~ "rejuvenation piece",
      `core-fragment` == "pyramid" ~ "pyramid",
      `core-fragment` == "pyramid core" ~ "pyramid",
      `core-fragment` == "pyramid-unidirectional" ~ "pyramid",
      `core-fragment` == "core fragment-pyramid" ~ "pyramid",
      `core-fragment` == "cylinder" ~ "cylinder/prismatic",
      `core-fragment` == "prismatic" ~ "cylinder/prismatic",
      `core-fragment` == "shapeless" ~ "cylinder/prismatic",
      `core-fragment` == "multidirectional core fragment" ~ "shapeless",
      `core-fragment` == "multidirectional flake core" ~ "shapeless",
      `core-fragment` == "multidirectional bladelet core" ~ "shapeless",
      `core-fragment` == "mixed core" ~ "shapeless",
      `core-fragment` == "mixed-percussion" ~ "shapeless",
      `core-fragment` == "mixed" ~ "shapeless",
      `core-fragment` == ("bipolar-percussion") ~ "bipolar",
      `core-fragment` == "flat" ~ "flat",
      `core-fragment` == "core on flake" ~ "core on flake",
      `core-fragment` == "bullet" ~ "bullet",
      TRUE ~ `core-fragment`

    )
  )  %>%
  mutate(
    core = case_when(
      `core-fragment` %in% c(
        "pyramid",
        "pyramid core",
        "pyramid-unidirectional",
        "core fragment-pyramid",
        "cylinder/prismatic",
        "shapeless",
        "multidirectional core fragment",
        "multidirectional flake core",
        "multidirectional bladelet core",
        "mixed core",
        "mixed-percussion",
        "mixed",
        "bipolar-percussion",
        "flat",
        "core on flake",
        "bullet"
      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    ),
    corefrag = case_when(
      `core-fragment` %in% c(
        "rejuvenation piece",
        "rejuvention piece-side",
        "core preparation piece-platform",
        "core preparation",
        "core preparation piece",
        "core preperation",
        "crested bladelet" ,
        "primary flake" ,
        "primary blade"
      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    )
  ) %>%
  # Convert empty strings to NA in `core-fragment` column
  mutate(`core-fragment` = if_else(`core-fragment` == "", NA_character_, `core-fragment`)) %>%
  # Convert "pyramid-percussion" to "pyramid" in `core-fragment` column
  mutate(
    `core-fragment` =
      if_else(
        `core-fragment` == "pyramid-percussion",
        "pyramid",
        `core-fragment`
      )
  ) %>%
  # Update existing `core` column with selected values from `core-fragment` column
  mutate(core = if_else(
    `core-fragment` %in%
      c(
        "pyramid",
        "cylinder/prismatic",
        "flat",
        "bullet",
        "core on flake",
        "pyramid"
      ),
    `core-fragment`,
    core
  )) %>%
  # Add 'core tablet' and 'core fragment' from `core-fragment` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-fragment` == "core tablet" ~ "core tablet",
      `core-fragment` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-typology` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology`
  mutate(`core-typology` = if_else(
    `core-typology` %in% c("rejuvenation piece", "core fragment"),
    NA_character_,
    `core-typology`
  ))  %>% 
# Update the 'blank' column based on conditions in the 'core-fragment' column
  mutate(blank = if_else(
    `core-fragment` %in% c(
      "core fragment",
      "core tablet",
      "rejuvenation piece",
      "pyramid",
      "crested bladelet",
      "shapeless",
      "primary flake",
      "flat",
      "bullet",
      "core on flake",
      "cylinder/prismatic",
      "primary blade",
      "rejuvenation piece-side",
      "core platform rejuvention"
    ),
    NA_character_,
    blank
  ))

# take a look
# rev(sort(table(tl_final$core)))
# rev(sort(table(tl_final$corefrag)))
# rev(sort(table(tl_final$`core-fragment`)))
# rev(sort(table(tl_final$blank)))
```

@@

```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
    `Cores (n)` = sum(!is.na(core) & core != ""),
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) &
                                            corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

@@

tl_final_excavation_data_volume_artefact_counts <- 
tl_final_excavation_data_volume %>% 
  group_by(label2) %>% 
   summarise(
    blade_count =              sum(blank == "blade", na.rm = TRUE),
    flake_count =              sum(blank == "flake", na.rm = TRUE),
    bladelet_count =           sum(blank == "bladelet", na.rm = TRUE),
    tool_count =               sum(retouch == 1, 
                                   na.rm = TRUE) + sum(utilization == 1, 
                                                       na.rm = TRUE),
    core_count =               sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count =  sum(!is.na(`corefrag`), 
                                   na.rm = TRUE),
    retouched_tool_count =     sum(retouch == 1, 
                                   na.rm = TRUE),
    utilized_tool_count =      sum(utilization == 1, 
                                   na.rm = TRUE),
    shine_tool_count =         sum(`sickle shine` == 1, 
                                   na.rm = TRUE),
    tool_type_count = n_distinct(tooltypegroup),
    excavation_volume = unique(excavation_volume)

  ) %>% 
 # BM:  "total lithics" is everything except sickle shine, 
 # "retouch frequency" is retouched_tool_count / "total lithics"
  rowwise() %>% 
  mutate(total_lithics = sum(blade_count,
                             flake_count,
                             bladelet_count,
                             tool_count,
                             core_count,
                             core_rejuvenation_count,
                             retouched_tool_count,
                             utilized_tool_count)) %>% 
  mutate(retouch_frequency =  retouched_tool_count / total_lithics)

>>>>

tl4 <- tl4 %>%
  mutate(
    `Core Typology` = str_to_lower(`Core Typology`),
    `Core Typology` = str_trim(`Core Typology`),
    # BM: what does this regex do in the next line?
    `Core Typology` = str_replace_all(`Core Typology`, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(`Core Typology` = case_when(
    `Core Typology` == "flatcore" ~ "flat-pressure",
             `Core Typology` == "rejuvention" ~ "rejuvenation piece-NA",
             `Core Typology` == "pressureprymidal core" ~ "pyramid-pressure",
             `Core Typology` == "pressureflat core" ~ "flat-pressure",
             `Core Typology` == "pressurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prussurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prismaticcore" ~ "cylinder/prismatic-percussion",
             `Core Typology` == "pyramidalpressure core" ~ "pyramid-pressure",
             `Core Typology` == "flatcore with one debitage surface/pressure" ~ "flat-pressure",
             `Core Typology` == "corefragment" ~ "core fragment-NA",
             `Core Typology` == "cilandrical/pressure" ~ "cylinder/prismatic-pressure",
             `Core Typology` == "bullet/pressure" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/pressure" ~ "pyramid-pressure",
             `Core Typology` == "percussion/pyramidal" ~ "pyramid-pressure",
             `Core Typology` == "flat/pressure" ~ "flat-pressure",
             `Core Typology` == "pyramidal/ percussion??" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ peressure" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ pressure" ~ "pyramid-pressure",
             `Core Typology` == "pressure" ~ "pyramid-pressure",
             `Core Typology` == "multidirectional/percussion" ~ "shapeless-percussion",
             `Core Typology` == "prismatic/ percussion" ~ "pyramid-pressure",
             `Core Typology` == "flatcore/ unidirectional/pressure" ~ "flat-pressure",
             `Core Typology` == "heavilyused" ~ "NA",
             `Core Typology` == "burnt" ~ "NA",
             `Core Typology` == "patinated" ~ "NA",
             `Core Typology` == "pyramidal(bullet)/pressure" ~ "bullet-pressure",
             `Core Typology` == "cylandrycal/ bidirectional pressure core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure bullet core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "bladelet" ~ "bullet-pressure",
             `Core Typology` == "lip" ~ "NA",
             `Core Typology` == "bladlet" ~ "bullet-pressure",
             `Core Typology` == "bladelet(30.56)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(28.55.6)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(16.28.1)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(40 5.7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(23.2 2.3)" ~ "bullet-pressure",
             `Core Typology` == "bladelet-flakeblade(41 16)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(3512.6/7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(27.307.5)" ~ "bullet-pressure",
    `Core Typology` == "pressure/pyramidal" ~ "pyramid-pressure",
    `Core Typology` == "bladeletcore" ~ "bullet-pressure",
    `Core Typology` == "pyramidal? / pressure" ~ "pyramid-pressure",
    `Core Typology` == "prismatic" ~ "pyramid-pressure",

    `Core Typology` == "0" ~ "",
    `Core Typology` == "na" ~ "",
    `Core Typology` == "na" ~ "",

@@

      `Typology` == "" ~ "",
      `Typology` == "na" ~ "",

@@

```{r}
tl_final <- tl_final %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`))  %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`)) %>%
  mutate(
    `core-fragment` = case_when(
      `core-fragment` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-fragment` == "rejuvention piece-side" ~ "rejuvenation piece",
      `core-fragment` == "rejuvenation piece-side" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece-platform" ~ "core tablet",
      `core-fragment` == "core preparation piece-platform" ~ "core tablet",
      `core-fragment` == "core platform rejuvention" ~ "core tablet",
      `core-fragment` == "core platform rejuvention " ~ "core tablet",
      `core-fragment` == "core preparation" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece" ~ "rejuvenation piece",
      `core-fragment` == "core preperation" ~ "rejuvenation piece",
      `core-fragment` == "pyramid" ~ "pyramid",
      `core-fragment` == "pyramid core" ~ "pyramid",
      `core-fragment` == "pyramid-unidirectional" ~ "pyramid",
      `core-fragment` == "core fragment-pyramid" ~ "pyramid",
      `core-fragment` == "cylinder" ~ "cylinder/prismatic",
      `core-fragment` == "prismatic" ~ "cylinder/prismatic",
      `core-fragment` == "shapeless" ~ "shapeless",
      `core-fragment` == "multidirectional core fragment" ~ "shapeless",
      `core-fragment` == "multidirectional flake core" ~ "shapeless",
      `core-fragment` == "multidirectional bladelet core" ~ "shapeless",
      `core-fragment` == "mixed core" ~ "shapeless",
      `core-fragment` == "mixed-percussion" ~ "shapeless",
      `core-fragment` == "mixed" ~ "shapeless",
      `core-fragment` == ("bipolar-percussion") ~ "bipolar",
      `core-fragment` == "flat" ~ "flat",
      `core-fragment` == "core on flake" ~ "core on flake",
      `core-fragment` == "bullet" ~ "bullet",
      `core-fragment` == "NA" ~ "",
      TRUE ~ `core-fragment`

    )
  )  %>%
  mutate(
    core = case_when(
      `core-fragment` %in% c(
        "pyramid",
        "cylinder/prismatic",
        "core fragment",
        "cylinder/prismatic",
        "bullet",
        "flat",
        "core on flake",
        "bipolar",
        "mixed core",
        "mixed-percussion",
        "mixed",
        "bipolar-percussion",
        "flat",
        "core on flake",
        "bullet",
        "pyramid",
        "flat",
        "cylinder/prismatic ",
        "bipolar ",
        "shapeless ",
        "bullet "

      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    ),
    corefrag = case_when(
      `core-fragment` %in% c(
        "rejuvenation piece",
        "core tablet",
        "primary flake",
        "primary blade",
        "crested bladelet",
        "core preperation",
        "crested bladelet" ,
        "primary flake" ,
        "primary blade",
        "core fragment "

      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    )
  ) %>%
  # Convert empty strings to NA in `core-fragment` column
  mutate(`core-fragment` = if_else(`core-fragment` == "", NA_character_, `core-fragment`)) %>%
  # Convert "pyramid-percussion" to "pyramid" in `core-fragment` column
  mutate(
    `core-fragment` =
      if_else(
        `core-fragment` == "pyramid-percussion",
        "pyramid",
        `core-fragment`
      )
  ) %>%
  # Update existing `core` column with selected values from `core-fragment` column
  mutate(core = if_else(
    `core-fragment` %in%
      c(
        "shapeless",
        "pyramid",
        "cylinder/prismatic",
        "flat",
        "bullet",
        "core on flake",
        "pyramid"
      ),
    `core-fragment`,
    core
  )) %>%
  # Add 'core tablet' and 'core fragment' from `core-fragment` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-fragment` == "core tablet" ~ "core tablet",
      `core-fragment` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-typology` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology`
  mutate(`core-typology` = if_else(
    `core-typology` %in% c("rejuvenation piece", "core fragment"),
    NA_character_,
    `core-typology`
  ))  %>% 
# Update the 'blank' column based on conditions in the 'core-fragment' column
  mutate(blank = if_else(
    `core-fragment` %in% c(
      "core fragment",
      "core tablet",
      "rejuvenation piece",
      "pyramid",
      "crested bladelet",
      "shapeless",
      "primary flake",
      "flat",
      "bullet",
      "core on flake",
      "cylinder/prismatic",
      "primary blade",
      "rejuvenation piece-side",
      "core platform rejuvention"
    ),
    NA_character_,
    blank
  ))

# take a look
 rev(sort(table(tl_final$core)))
 rev(sort(table(tl_final$corefrag)))
 rev(sort(table(tl_final$`core-fragment`)))
 rev(sort(table(tl_final$`core-typology`)))
 rev(sort(table(tl_final$blank)))
```

@@

```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
   `Cores (n)` = sum((!is.na(`core-typology`) & `core-typology` != "") |
                      (!is.na(core) & core != ""), na.rm = TRUE),  # Summing variables from both columns
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) &
                                            corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Manually calculate the totals for each area and grand total
total_d1 <- sum(summary_table_tool_area$d1, na.rm = TRUE)
total_ta <- sum(summary_table_tool_area$ta, na.rm = TRUE)
total_tp1 <- sum(summary_table_tool_area$tp1, na.rm = TRUE)
grand_total <- total_d1 + total_ta + total_tp1

# Calculate the percentages for each cell based on the total count
summary_table_tool_area_percentage <- summary_table_tool_area %>%
  mutate(
    percent_d1 = round(d1 / total_d1 * 100, 1),
    percent_ta = round(ta / total_ta * 100, 1),
    percent_tp1 = round(tp1 / total_tp1 * 100, 1)
  ) %>%
  # Add the total_percent column, calculating based on grand total
  mutate(total_percent = round((d1 + ta + tp1) / grand_total * 100, 1))

# Select only the necessary columns (tooltypegroup, the percentage columns, and total_percent)
summary_table_tool_area_percentage <- summary_table_tool_area_percentage %>%
  select(tooltypegroup, percent_d1, percent_ta, percent_tp1, total_percent)

# write the percentage table to a CSV file
write_csv(summary_table_tool_area_percentage,
          here("data/derived-data/tbl-002-summary_table_tool_area_percentage.csv"))

kable(
  summary_table_tool_area_percentage,
  col.names = c("Tools", "D1 (%)", "TA (%)", "TP1 (%)", "Total (%)"),
  caption = "Tool Typology in the three main areas (Percentage)"
)

@@

tl_final_excavation_data_volume_artefact_counts <- 
tl_final_excavation_data_volume %>% 
  group_by(label2) %>% 
   summarise(
    blade_count =              sum(blank == "blade", na.rm = TRUE),
    flake_count =              sum(blank == "flake", na.rm = TRUE),
    bladelet_count =           sum(blank == "bladelet", na.rm = TRUE),
    tool_count =               sum(retouch == 1, 
                                   na.rm = TRUE) + sum(utilization == 1, 
                                                       na.rm = TRUE),
    core_count =                sum(!is.na(`core-typology`), na.rm = TRUE) + sum(!is.na(core), na.rm = TRUE),

    core_rejuvenation_count =  sum(!is.na(`corefrag`), 
                                   na.rm = TRUE),
    retouched_tool_count =     sum(retouch == 1, 
                                   na.rm = TRUE),
    utilized_tool_count =      sum(utilization == 1, 
                                   na.rm = TRUE),
    shine_tool_count =         sum(`sickle shine` == 1, 
                                   na.rm = TRUE),
    tool_type_count = n_distinct(tooltypegroup),
    excavation_volume = unique(excavation_volume)

  ) %>% 
 # BM:  "total lithics" is everything except sickle shine, 
 # "retouch frequency" is retouched_tool_count / "total lithics"
# SS: I think "total lithics" would be the total of blade, flake,bladelet, core and core rejuvenation counts (as we already cound the tools in bladelet, blade, or flake) 
  rowwise() %>% 
  mutate(total_lithics = sum(blade_count,
                             flake_count,
                             bladelet_count,
                             core_count,
                             core_rejuvenation_count

                             )) %>% 
  mutate(retouch_frequency =  retouched_tool_count / total_lithics)
